The impacts of demography on linked selection in maize

### 2015-07-01 manuscript.tex
comments

\section{Results}
\subsection{Patterns of variability differ between genic and
  nongenic regions of the genome}
Previous research of the maize domestication process has relied upon
observations drawn from genic DNA \textcolor{red}{(several references
  here)}. Our data were generated through whole genome sequencing,
which eliminated this constraint. Importantly, we observed substantial
differences in patterns of diversity between genic and non-genic regions
of the genome for both maize and teosinte. For maize, mean pairwise
diversity ($\pi$) within genes was significantly lower than at
positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
pattern of significantly greater diversity outside of genes was observed in
teosinte, but to a larger extent. For teosinte we observed $\pi$
within genes to be 0.0088 and at least 5kb from genes it was 0.1153
(p$\approx$0). These obserrvations suggest that genes are not evolving
neutrally in maize or teosinte. Instead, some form of selection is likely
reducing diversity within genes. Additionally, these observations
suggest that demographic inference, which relies on the assumption of
neutrally evolving DNA, will be more accurate if it is based on
observations from non-genic genetic material.

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous and synonymous substitutions in maize and
teosinte since divergence from \emph{Tripsacum}, roughly \textcolor{red}{ZZZ} years before
present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure
\ref{fig:hardSweeps}). Additionally, diversity around maize substitutions
not seen in teosinte was investigated. These sites have the potential
to correspond to recent maize sweeps, occuring after the split from
teosinte. Again, no difference was observed between diversity around
synonymous and nonsynonymous substitutions
\textcolor{red}{(Supplemental figure here)}. Together, these
observations suggest that hard sweeps are not a primary form of
selection for either maize or teosinte.

@@

\begin{materials}
\subsection{Plant materials}
Accessions studied were selected from the Maize HapMap2
panel \cite{chia2012} . Principal component analysis was employed to ensure that
closely related individuals were not included due to their potential
to bias results \textcolor{red}{(maybe a supplemental figure here)}. Ultimately, 23 maize inbreds derived from a diverse
assortment of landraces were selected for inclusion. Thirteen teosinte
inbred lines, all members of the subspecies Z. \emph{mays}
ssp. \emph{parviglumis}, were utilized. Sequences were mapped to the
maize B73 version 3 reference genome \cite{schnable2009}
(ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).

\subsection{Interpolating genetic position}
For many of the following analyses, physical position along a
chromosome was a less relevant measure of map location than was genetic
position. Therefore, physical positions were converted to genetic
positions by interpolating from the NAM genetic map (REFERENCE), which
provides a 1 cM resolution for physical to genetic conversion. Within
R \cite{R2014}, physical positions with corresponding genetic positions in
the NAM map were used as anchors. Physical positions in our dataset
wihout corresponding genetic position were assigned genetic positions by scaling
the anchored genetic positions according to the physical distance
between the unlabeled position and the flanking anchors.

>>>>

\section{Results}
\subsection{Patterns of variability differ between genic and
  nongenic regions of the genome}
Previous research of the maize domestication process has relied upon
observations drawn from genic DNA \textcolor{red}{(several references
  here)}. Our data were generated through whole genome sequencing,
which eliminated this constraint. Importantly, we observed substantial
differences in patterns of diversity between genic and non-genic regions
of the genome for both maize and teosinte. For maize, mean pairwise
diversity ($\pi$) within genes was significantly lower than at
positions at least 5kb away from genes (0.00668 within, 0.00691 away, p$<$2e-44). The same
pattern of significantly greater diversity outside of genes was observed in
teosinte, but to a larger extent. For teosinte we observed $\pi$
within genes to be 0.0088 and at least 5kb from genes it was 0.1153
(p$\approx$0). \jri{that number is impossible, check again. off by maybe an order of magnitude!}
These observations suggest that genes are not evolving
neutrally in maize or teosinte. Instead, some form of selection is likely
reducing diversity within genes. Additionally, these observations
suggest that demographic inference, which relies on the assumption of
neutrally evolving DNA, will be more accurate if it is based on
observations from non-genic genetic material.

\subsection{Hard sweeps do not shape maize (or teosinte) diversity}
A mutation that is immediately beneficial and positively selected leaves a classical hard
sweep signature in the genome, whereby genetic diversity surrounding
that mutation is reduced as the haplotype where the mutation first
arose increases in frequency to fixation. The prevalence of such hard
sweeps was evaluated by comparing diversity
surrounding non-synonymous and synonymous substitutions in maize and
teosinte since divergence from \emph{Tripsacum}, roughly \textcolor{red}{ZZZ} years before
present \textcolor{red}{(ref here)}. For both taxa, no difference in diversity
surrounding these classes of substitutions was identified (Figure
\ref{fig:hardSweeps}). \jri{weird formatting that the first figure referenced is fig. 2. something about PNAS style and methods?}Additionally, diversity around maize substitutions not seen in teosinte was investigated. 
These sites have the potential
to correspond to recent maize sweeps, occuring after the split from
teosinte. Again, no difference was observed between diversity around
synonymous and nonsynonymous substitutions
\textcolor{red}{(Supplemental figure here)}. Together, these
observations suggest that hard sweeps are not a primary form of
selection for either maize or teosinte. \jri{add Enard citation and GERP results (supp. figures)}

@@

\begin{materials}

\jri{either we need to provide all scripts in a single github repo with a readme, or provide links to individual scripts in the methods as we go along}

\subsection{Plant materials}
Accessions studied were selected from the Maize HapMap2
panel \cite{chia2012}. \jri{aren't you using the high coverage teosinte? if so I think you need to cite Lemmon 2014 Plos Gen} 
Principal component analysis was employed to ensure that closely related individuals were not included due to their potential to bias results \textcolor{red}{(maybe a supplemental figure here)}. \jri{yes supp. figure good idea}
Ultimately, 23 maize inbreds derived from a diverse assortment of landraces were selected for inclusion. 
Thirteen teosinte inbred lines, all members of  subspecies Z. \emph{mays}
ssp. \emph{parviglumis}, were utilized. 
Sequences were mapped to the maize B73 version 3 reference genome \cite{schnable2009} \url(ftp://ftp.ensemblgenomes.org/pub/plants/release-22/fasta/zea\_mays/dna/).

\subsection{Interpolating genetic position}
For many of the following analyses, physical position along a chromosome was a less relevant measure of map location than was genetic position. 
Therefore, physical positions were converted to genetic positions by interpolating from the NAM genetic map \citex \jri{Wallace 2014 no?}, which provides a 1 cM resolution for physical to genetic conversion. 
Within R \cite{R2014}, physical positions with corresponding genetic positions in the NAM map were used as anchors. 
Physical positions in our dataset wihout corresponding genetic position were assigned genetic positions by scaling the anchored genetic positions according to the physical distance between the unlabeled position and the flanking anchors.
